\chapter{Group 16 Evaluation} \label{cap:tests}

As part of Cloudy Day, our group evaluated Group 16's project, the \textit{European Country Metrics Aggregator}. The goal of this evaluation was to validate the public API from the perspective of an external team, checking both functional correctness and behaviour under load. Each group member tested one endpoint with a smoke test and another endpoint with a stress test.

The smoke tests focused on \textbf{API contract validation}: status codes, JSON structure, required fields, error handling and relevant edge cases. The stress tests used Locust to generate concurrent traffic and identify the point where latency or failures became significant. For stress scenarios, a response was only considered successful when both the HTTP response and the response body matched the expected contract within the configured latency threshold.

\section{Test Scope}

The evaluation covered eight endpoints across the location, authentication and forum areas of the API. Table~\ref{tab:g16_scope} summarizes the responsibilities assigned to each tester.

\begin{table}[H]
    \centering
    \caption{Group 16 evaluation scope}
    \label{tab:g16_scope}
    \begin{tabular}{p{0.30\textwidth} p{0.20\textwidth} p{0.40\textwidth}}
        \hline
        \textbf{Tester} & \textbf{Test type} & \textbf{Endpoint} \\\hline
        Francisco Encarnação & Smoke & \texttt{POST /forum/posts} \\
        Francisco Encarnação & Stress & \texttt{GET /forum/posts} \\
        Daniel Nunes & Smoke & \texttt{GET /locations/\{id\}} \\
        Daniel Nunes & Stress & \texttt{GET /locations/code/\{code\}} \\
        Daniel Sousa & Smoke & \texttt{GET /locations} \\
        Daniel Sousa & Stress & \texttt{GET /locations/compare} \\
        Daniel Carvalho & Smoke & \texttt{POST /auth/login} \\
        Daniel Carvalho & Stress & \texttt{GET /locations/top} \\\hline
    \end{tabular}
\end{table}

\newpage 

\section{Smoke Test Results}

The smoke tests showed that most basic API behaviours were implemented correctly. The location endpoints were the most consistent: listing locations and retrieving a location by identifier returned the expected structures, handled missing resources with \texttt{404 Not Found}, rejected invalid identifiers with \texttt{400 Bad Request}, and supported the documented year filtering behaviour.

The forum and authentication endpoints revealed more important issues. In \texttt{POST /forum/posts}, a valid request using the field name documented in the specification, \texttt{authorID}, was rejected. The same request succeeded when using \texttt{authorId}, showing a \textbf{case-sensitivity mismatch} between the specification and the implementation. The endpoint also allowed creating posts without authentication, which means a client could create posts on behalf of another user by guessing an identifier, which poses a \textbf{severe security risk}.

For \texttt{POST /auth/login}, most normal and negative paths behaved correctly: registration succeeded, duplicate registration returned \texttt{409 Conflict}, valid login returned \texttt{200 OK}, unknown credentials returned \texttt{401 Unauthorized}, and missing fields returned \texttt{400 Bad Request}. However, the successful login response did not include the expected \texttt{User} object, a request with \texttt{Content-Type: text/plain} returned \texttt{500 Internal Server Error}, and the cleanup request used by the test returned \texttt{404 Not Found}.

\begin{table}[H]
    \centering
    \caption{Smoke test summary}
    \label{tab:g16_smoke_summary}
    \begin{tabular}{p{0.25\textwidth} p{0.14\textwidth} p{0.50\textwidth}}
    \hline
    \textbf{Endpoint} & \textbf{Result} & \textbf{Main observations} \\\hline
    \texttt{POST /forum/posts} & Partial pass & Missing fields were validated, but the specification used \texttt{authorID} while the API expected \texttt{authorId}. \textbf{Post creation did not require authentication.} \\
    \texttt{GET /locations/\{id\}} & Pass & Valid IDs, missing resources, invalid IDs, year filters, path variants and unsupported methods were handled correctly. \\
    \texttt{GET /locations} & Pass & 20 requests completed with 0 failures, an average response time of \texttt{62.7ms}, and a maximum response time of \texttt{231.2ms}. \\
    \texttt{POST /auth/login} & Partial pass & 13 out of 16 checks passed. The response schema, wrong content-type handling and cleanup route had issues. \\\hline
    \end{tabular}
\end{table}

\newpage 

\section{Stress Test Results}

The stress tests showed two different behaviours. Some endpoints remained functionally correct but \textbf{degraded under high concurrency}, while \texttt{GET /locations/top} handled the tested load profile without failures.

The \texttt{GET /forum/posts} stress test had to be split into two phases. In the first phase, the unfiltered endpoint failed every request: all \texttt{17,931} requests to the global endpoint returned failures, even though the \texttt{userId} query parameter is documented as optional. In the second phase, the filtered endpoint \texttt{GET /forum/posts?userId=X} behaved correctly at lower load but reached a \texttt{7.34\%} failure rate during the peak, with a maximum response time of \texttt{17,740ms}.

The location stress tests were more stable in terms of correctness. \texttt{GET /locations/code/\{code\}} returned valid data for most requests but showed clear saturation under the highest load stages, with \texttt{5.72\%} failures and a \texttt{p95} response time of \texttt{3700ms}. In contrast, \texttt{GET /locations/top} completed \texttt{25,502} requests with no failures and had a \texttt{p95} response time at \texttt{190ms}, although its maximum response time of \texttt{2972ms} was close to the configured threshold.

\begin{table}[H]
    \centering
    \caption{Stress test summary}
    \label{tab:g16_stress_summary}
    \begin{tabular}{p{0.34\textwidth} p{0.12\textwidth} p{0.11\textwidth} p{0.32\textwidth}}
    \hline
    \textbf{Endpoint} & \textbf{Requests} & \textbf{Failures} & \textbf{Latency notes} \\\hline
    \texttt{GET /forum/posts} & \texttt{17,931} & \texttt{100\%} & Global route failed when \texttt{userId} was omitted. \\
    \texttt{GET /forum/posts?userId=X} & \texttt{18,748} & \texttt{7.34\%} & Maximum response time reached \texttt{17,740ms}. \\
    \texttt{GET /locations/code/\{code\}} & \texttt{24,762} & \texttt{5.72\%} & \texttt{p95} reached \texttt{3700ms}; maximum was \texttt{21,486.7ms}. \\
    \texttt{GET /locations/compare} & \texttt{18,858} & \texttt{5.66\%} & \texttt{p95} reached \texttt{3600ms}; maximum was \texttt{22,546.9ms}. \\
    \texttt{GET /locations/top} & \texttt{25,502} & \texttt{0\%} & \texttt{p95} was \texttt{190ms}; maximum was \texttt{2972ms}. \\\hline
    \end{tabular}
\end{table}

\section{Conclusion and Recommendations}

Overall, the Group 16 API passed several important functional checks, especially for the location endpoints, but upon further inspection, the evaluation uncovered \textbf{contract mismatches}, \textbf{severe security gaps}, unsafe malformed-request handling and performance degradation under high concurrency. The recommendations include:

\begin{itemize}
    \item The API specification and implementation should be aligned and no mismatches should exist. Namely between \texttt{authorID} and \texttt{authorId} causing a valid documented request to fail and the login response schema should either include the expected \texttt{User} object or the contract should be updated to describe the actual response.
    \item Malformed requests should never produce internal server errors. The authentication endpoint returned \texttt{500 Internal Server Error} for an unsupported content type, which should instead be handled as a client error such as \texttt{400 Bad Request} or \texttt{415 Unsupported Media Type}. The global forum listing endpoint also returned failures when the optional \texttt{userId} parameter was omitted, so the backend logic should handle the documented optional case safely.
    \item The forum post creation endpoint needs authentication and authorization. Allowing unauthenticated clients to create posts using arbitrary user identifiers is a security risk because users can impersonate others.
    \item Finally, the endpoints that degraded under stress should be profiled at the data access layer. Reviewing database indexes, optimizing expensive queries and adding caching for repeated read-heavy requests (if not done already) can help improve performance.
\end{itemize}
